---
title: "Wisconsin Districts"
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false

library(mdsr)
library(sf)
library(mapproj)
library(maps)
library(ggspatial)
library(ggdark)
library(prettymapr)
library(raster)
library(fec16)
library(tidyverse)
library(conflicted)

conflict_prefer("filter","dplyr")
conflict_prefer("select", "dplyr")
```

```{r}
#| message: false
#| warning: false
district_elect <- fec16::results_house |>
  select(-runoff_votes, -runoff_percent, -footnotes) |>
  mutate(district = parse_number(district_id)) |>
  group_by(state, district) |>
  summarize(
    N = n(), 
    total_votes = sum(general_votes, na.rm = TRUE),
    d_votes = sum(ifelse(party == "DEM", general_votes, 0), na.rm = TRUE),
    r_votes = sum(ifelse(party == "REP", general_votes, 0), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    other_votes = total_votes - d_votes - r_votes,
    r_prop = r_votes / total_votes,  
    winner = ifelse(r_votes > d_votes, "Republican", "Democrat")
  )
wi_results <- district_elect |>
  filter(state == "WI")

wi_d3 <- fec16::results_house |>
  select(-runoff_votes, -runoff_percent, -footnotes) |>
  filter(state == "WI") |>
  filter(district_id == "03")

wi_d3
```

Wisconsin District 3 had 2 democrats running for the house seat.

```{r}
#| message: false
#| warning: false
#| include: false

# url <- "https://jadenchant.github.io/data/districts113.zip"
# districts <- st_read(paste0("/vsizip/vsicurl/", url))


# wi_shp <- districts |>
#   filter(STATENAME == "Wisconsin")

# st_write(wi_shp, "~/data/wisconsin_districts.gpkg", delete_dsn = TRUE)
```


```{r}
#| message: false
#| warning: false

districts <- st_read("https://jadenchant.github.io/data/wisconsin_districts.gpkg")

wi_merged <- districts |>
  st_transform(4326) |>
  inner_join(wi_results, by = c("DISTRICT" = "district"))

wi <- ggplot(data = wi_merged, aes(fill = winner)) +
  annotation_map_tile(zoom = 7, type = "cartodark", progress = "none") + 
  geom_sf(alpha = 0.4, color = "white", size = 0.2) +
  scale_fill_manual("Winner", values = c("Democrat" = "#2E5B88", "Republican" = "#8B1A1A")) + 
  geom_sf_text(aes(label = DISTRICT), color = "white", fontface = "bold", check_overlap = TRUE) + 
  ggtitle("Wisconsin House Winners by District") +
  labs(x = NULL, y = NULL) +
  dark_theme_gray(base_size = 12) + 
  theme(
    axis.ticks = element_blank(), 
    axis.text = element_blank(), 
    panel.grid = element_blank(),
    plot.margin = unit(c(0.2, 0.7, 0, 1.5), "cm")
  )

wi

wi_results |>                  
  select(-state)
```

The map shows the winners of the 2016 House elections in Wisconsin. The district shapes are very irregular. District 4 is extremely small when compared to district 7. There are 2 districts where there isn't even a republican running. The races where republicans have won the district, they have won by a small margin. However, the margin is not as small as the North Carolina districts that we looked at in class.
